A new form of transcorrelated Hamiltonian inspired by range-separated DFT

### 2020-11-19 cusp_j/paper/paper.tex
modified paper

The paper is organized as follows. 
In the first section, we derive the analytical form of a new Jastrow factor and the corresponding transcorrelated Hamiltonian. 
The, we perform a detailed numerical study of the behaviour of the solution of $\tilde{H}[\mu]$ on the helium atom: 
in the first part we analyze the speed of convergence of the ground state energy with respect to the basis set and the parameter $\mu$, 
then we study the behaviour of the right eigenvectors in real space and compare it with the exact ground state wave function. 
We show that the rapid convergence of the total energies coincides with a rapid convergence of the right eigenvectors in real space. 
Based on such encouraging results, we study the helium isoelectronic series from Li$^+$ to Ne$^{8+}$ and compare different schemes to obtain 
a value of $\mu$ which depends on the system and still provides a fast convergence of the total energies. 
Eventually, we study the H$_2$ molecule and compare different regimes of correlation.

\section{A new form of Jastrow factor for transcorrelated Hamiltonian}
This section is dedicated to the derivation of a new form of Jastrow factor mimicking at short range the correlation effects of RS-DFT. 
For the sake of simplicity and to remain as close as possible to the physical picture of the electronic correlation problem, 
we derive the equations of the transcorrelated Hamiltonian for the helium atom for a general Jastrow factor depending on the inter electronic distance.  
Thanks to the simplicity of the equations, we focus on the behaviour at small $r_{12}$ of different flavours of Hamiltonians 
and the corresponding discontinuity in their eigenvectors. 
We browse three different types of Hamiltonians: the usual physical Hamiltonian, the transcorrelated Hamiltonian together with the effective Hamiltonian used in RS-DFT.
Then, we derive the general analytical form of the corresponding new transcorrelated Hamiltonians $\tilde{H}[\mu]$, and its dependence as function of the parameter $\mu$.

\subsection{A simple physical picture: the case of the helium atom}
\subsubsection{The transcorrelated Hamiltonian for helium atom}
Let us write the Hamiltonian of the He atom in the ${\bf r}_{12} = \big|{\bf r}_1 - {\bf r}_2 \big|$ coordinate 
\begin{equation}
 H  = h_c + \frac{1}{r_{12}},
\end{equation}
where 
\begin{equation}
 \begin{aligned}
 \label{def:h_c}
 h_c = &-\frac{1}{2} \sum_{i=1}^2 \bigg(\deriv{}{r_i}{2} + \frac{2}{r_i} \deriv{}{r_i}{} + \frac{2 Z}{r_i}\bigg) \\
     &-\bigg( \deriv{}{r_{12}}{2} + \frac{2}{r_{12}} \deriv{}{r_{12}}{} \bigg) \\
     &-\bigg( \frac{\bd{r_1}}{r_1} \cdot \frac{\bd{r_{12}} }{r_{12}}  \deriv{}{r_1}{} + 
                \frac{\bd{r_2}}{r_2} \cdot \frac{\bd{r_{21}} }{r_{21}}  \deriv{}{r_2}{} \bigg),
 \end{aligned}
\end{equation}
\label{sec:he_j}
Now let us consider the Similarity transformed Hamiltonian $\tilde{H}[j]$
\begin{equation}
 \label{eq:ht_0}
 \begin{aligned}
 \tilde{H}[j]&= e^{-j(r_{12})} H e^{j(r_{12})},
 \end{aligned}
\end{equation}
where $j(r_{12})$ is a general Jastrow factor depending only on $r_{12}$. 
Therefore, the only new terms arising in $\tilde{H}[j]$ with respect to $H$ are those coming from the differential operator in $r_{12}$,
\begin{equation}
 \mathcal{T}[j] =  -e^{-j(r_{12})}\bigg( \deriv{}{r_{12}}{2} + \frac{2}{r_{12}} \deriv{}{r_{12}}{} \bigg)e^{j(r_{12})}.  
\end{equation}
Let us compute the action of $\mathcal{T}[j]$ on a general function $\phi(\br{}_1,\br{}_2)$ 
\begin{equation}
 \begin{aligned}
 \mathcal{T}[j]\phi(\br{}_1,\br{}_2) & =- e^{-j(r_{12})} \bigg[ \deriv{}{r_{12}}{2} + \frac{2}{r_{12}} \deriv{}{r_{12}}{}\bigg]  e^{j(r_{12})} \phi(\br{}_1,\br{}_2) \\ 
                            &=  -\deriv{\phi(\br{}_1,\br{}_2)}{r_{12}}{2} - \frac{2}{r_{12}} \deriv{\phi(\br{}_1,\br{}_2)}{r_{12}}{} \\
  &- 2 \deriv{j(r_{12})}{r_{12}}{} \deriv{\phi(\br{}_1,\br{}_2)}{r_{12}}{} \\ 
  &- \bigg[ \frac{2}{r_{12}} \deriv{j(r_{12})}{r_{12}}{}  + \deriv{j(r_{12})}{r_{12}}{2} + \bigg( \deriv{j(r_{12})}{r_{12}}{} \bigg)^2\bigg] \phi(\br{}_1,\br{}_2). 
 \end{aligned}
\end{equation}
Now, by defining the following operators 
\begin{equation}
 \label{eq:def_tt}
 \tilde{t}[j] = -2 \deriv{j(r_{12})}{r_{12}}{} \deriv{}{r_{12}}{},
\end{equation}
\begin{equation}
 \label{eq:def_wt}
 \tilde{W}[j] = -\frac{2}{r_{12}} \deriv{j(r_{12})}{r_{12}}{}  , 
\end{equation}
\begin{equation}
 \label{eq:def_wt}
 \tilde{w}[j] = -\deriv{j(r_{12})}{r_{12}}{2} - \bigg( \deriv{j(r_{12})}{r_{12}}{} \bigg)^2. 
\end{equation}
one can write the Similarity transformed Hamiltonian  as
\begin{equation}
 \label{eq:ht_phi}
 \tilde{H}[j] = h_c + \frac{1}{r_{12}}  + \tilde{W}[j] + \tilde{t}[j] + \tilde{w}[j] ,
\end{equation}
ore more explicitly 
\begin{equation}
 \tilde{H}[j] = h_c + \frac{1}{r_{12}}\bigg( 1 - 2 \deriv{j(r_{12})}{r_{12}}{}\bigg) + \tilde{t}[j] + \tilde{w}[j] .
\end{equation}
$\tilde{H}[j]$, because of the first-order differential operator in $\tilde{t}[j]$, is then non-hermitian,   
and the effective potential $ \frac{1}{r_{12}}  + \tilde{W}[j] + \tilde{w}[j]$ in $\tilde{H}[j]$ only depends on the shape of $j(r_{12})$.

\subsubsection{$r_{12} \rightarrow 0$ limits for different forms of Hamiltonians}
\label{sec:cusp}
Having established the general form of the transcorrelated Hamiltonian $\tilde{H}[j]$, in the case of the helium atom, one can then study the behaviour of the latter when $r_{12}\rightarrow 0$ and compare it to two other types of Hamiltonian: the usual physical Hamiltonian and that entering the RS-DFT framework. 
The analysis of the $r_{12}\rightarrow 0$ behaviour leads to the different conditions that the eigenvectors of such operators must fulfill. For the sake of simplicity of the notations we focus on the ground state of these operators but the arguments are general for any bounded eigenvectors.

In the case of the usual Hamiltonian, the exact ground state wave function $\psiex$ must satisfy the Schroedinger equation in real space 
\begin{equation}
 H \psiex(\br{}_1,\br{}_2) = E_0 \psiex(\br{}_1,\br{}_2)\quad \forall (\br{}_1,\br{}_2).
\end{equation}
When looking at $r_{12}\approx 0$, all terms multiplying $\frac{1}{r_{12}}$ in $H \psiex(\br{}_1,\br{}_2)$ must remain finite, which, according to Eq.\eqref{def:h_c}, translates into
\begin{equation}
 \label{eq:cusp_0}
 \lim_{r_{12}\rightarrow 0}\bigg( \frac{2}{r_{12}} \deriv{}{r_{12}}{} -\frac{1}{r_{12}}\bigg) \psiex(\bd{r_1},\bd{r_2})  = c,\quad c<\infty 
\end{equation}
or equivalently by multiplying by $r_{12}$
\begin{equation}
 \lim_{r_{12}\rightarrow 0}\bigg( 2 \deriv{\psiex(\bd{r_1},\bd{r_2})}{r_{12}}{} -\psiex(\bd{r_1},\bd{r_2}) \bigg) = 0, 
\end{equation}
which translates into the famous cusp condition for antiparallel spins of Kato\cite{Kat-CPAM-57}, 
\begin{equation}
 \deriv{\psiex(\bd{r_1},\bd{r_2})}{r_{12}}{}\Bigr|_{r_{12}=0} = \frac{1}{2} \psiex(r_{12}=0). 
\end{equation} 
For a pedagogical and general introduction to the cusp conditions, see Ref.\onlinecite{Tew-JCP-08}.

Regarding now the similarity transformed Hamiltonians, the exact ground state eigenvector must also satisfy an eigenvalue equation in real-space,
\begin{equation}
 \label{eq:hpi0_ex}
 \tilde{H}[j] \phiex(\br{}_1,\br{}_2) = E_0 \phiex(\br{}_1,\br{}_2) \quad \forall (\br{}_1,\br{}_2).
\end{equation}
Similarly to Eq. \eqref{eq:cusp_0}, when looking at $r_{12}\approx 0$, the terms multiplying $\frac{1}{r_{12}}$ in $\tilde{H}[j]\phiex(\br{}_1,\br{}_2)$ must remain finite, 
which involves also the term $\tilde{W}[j]$ in $\tilde{H}[j]$ (see Eq.\eqref{eq:def_wt} and \eqref{eq:ht_phi}. Therefore, imposing the finiteness of the limit $r_{12}\rightarrow 0$ of Eq.\eqref{eq:hpi0_ex} translates into
\begin{equation}
 \label{eq:cusp_phi_0}
 \lim_{r_{12}\rightarrow 0}\bigg( \frac{2}{r_{12}} \deriv{}{r_{12}}{} -\frac{1}{r_{12}}\bigg[ 1 - 2 \deriv{j(r_{12})}{r_{12}}{}\bigg]\bigg) \phiex(\bd{r_1},\bd{r_2})  = c,\quad c<\infty.
\end{equation}
Then, the term $\frac{1}{r_{12}}\bigg[ 1 - 2\deriv{j(r_{12})}{r_{12}}{}\bigg]$ in Eq.\eqref{eq:cusp_phi_0} looks like an effective electron electron interaction modified by the presence of the Jastrow factor in $\tilde{H}[j]$. 
Therefore, as long as one imposes that 
\begin{equation}
 \label{eq:cusp_phi_1}
  \deriv{j(r_{12})}{r_{12}}{}\bigg|_{r_{12}=0} = \frac{1}{2},
\end{equation}
which is nothing but the cusp condition for the Jastrow factor $j(r_{12})$, the scalar term proportional to $\frac{1}{r_{12}}$ in Eq.\eqref{eq:cusp_phi_0} vanishes at $r_{12}=0$, and then obtain a non-divergent effective electron-electron interaction. 
Within the condition of Eq.\eqref{eq:cusp_phi_1}, the general condition of Eq.\eqref{eq:cusp_phi_0} becomes 
\begin{equation}
 \label{eq:cusp_phi_2}
 \deriv{\phiex(\bd{r_1},\bd{r_2})}{r_{12}}{}\Bigr|_{r_{12}=0} = 0, 
\end{equation}
which implies that as long as the Jastrow factor contains the cusp conditions, the eigenfunction of $\tilde{H}[j]$ are cuspless. 
For instance, a simple Jastrow factor of the form $j(r_{12}) = \frac{1}{2} r_{12}$ release $\phiex(\br{}_1,\br{}_2)$ from 
the constraint of fulfilling the cusp condition.

Another form of effective Hamiltonians leading to cuspless eigenfunctions are those obtained from RS-DFT 
where the Coulomb interaction $1/r_{12}$ is split into a non-divergent long-range interaction $\text{erf}(\mu r_{12})/r_{12}$ and a complementary short-range interaction $\text{erfc}(\mu r_{12})/r_{12}$, where $\text{erf}(x)$ and $\text{erfc}(x)$ are the error and complementary error functions, respectively. 
The parameter which controls such a splitting is the so-called range separation parameter $\mu$:  RS-DFT reduces the usual WFT when $\mu \rightarrow \infty$, and $\mu \rightarrow 0$ gives back the usual Kohn-Sham theory.  
In practice, RS-DFT introduces a self-consistent Schroedinger-like equation which must be fulfilled by an effective wave function  $\Psi^\mu$.
Applied to a two electron system and looking in regions where $r_{12}\approx 0$, the leading terms of the self-consistent Schroedinger-like equation of RS-DFT reads 
\begin{equation}
 \label{eq:cusp_psi_mu_0}
 \lim_{r_{12}\rightarrow 0}\bigg( \frac{2}{r_{12}} \deriv{}{r_{12}}{} -\frac{\text{erf}(\mu r_{12})}{r_{12}}\bigg) \Psi^\mu(\bd{r_1},\bd{r_2})  = c,\quad c<\infty,  
\end{equation}
and as the long-range interaction is non divergent at $r_{12}=0$ as long as $\mu$ remains finite 
\begin{equation}
 \lim_{r_{12} \rightarrow 0} \frac{\text{erf}(\mu r_{12})}{r_{12}} = \frac{2 \mu}{\sqrt{\pi}} , 
\end{equation}
one obtains that 
\begin{equation}
 \label{eq:cusp_psi_mu_1}
 \deriv{\Psi^\mu(\bd{r_1},\bd{r_2})}{r_{12}}{}\Bigr|_{r_{12}=0} = 0, 
\end{equation}
which means that the wave functions $\Psi^\mu$, just as $\phiex$ are cuspless, because they deal with an effective non-divergent interaction.

@@

\subsection{A new form of Jastrow factor $j(\mu,r_{12})$ mimicking a long-range interaction}
In section \ref{sec:cusp} we reviewed the forms of limit conditions at $r_{12}=0$ involved in different eigenvalue equations, which of course depends on the nature of the interaction at $r_{12}=0$. 
In that section we derive a new Jastrow factor inspired by the limit condition of the similarity transformed Hamiltonian (see Eq.\eqref{eq:cusp_phi_0}) which mimics that of RS-DFT (see Eq.\eqref{eq:cusp_psi_mu_0}), at least at leading order. 
\subsubsection{The working equation for $j(\mu,r_{12})$ }
Eq. \eqref{eq:ht_phi} gives the similarity transformed Hamiltonian for an arbitrary Jastrow factor, 
and with respect to the usual Hamiltonian, $\tilde{H}[j]$ contains two new classes of terms: 
a new electron-electron potential $\tilde{w}[j]+ \tilde{W}[j]$ and a new differential operator  $\tilde{t}[j]$. 
Of course, these new terms depends on the explicit form of the Jastrow factor $j(r_{12})$ chosen. 
Now, we impose the form of $j(r_{12})$ such that it mimics, at leading order in $1/r_{12}$, the long-range effective interaction of RS-DFT. 
Mathematically, this conditions implies that 
\begin{equation}
 \begin{aligned}
 \label{def_j_00}
 \tilde{W}[j] + \frac{1}{r_{12}}&= \frac{\text{erf}(\mu r_{12})}{r_{12}} \\ 
\Leftrightarrow -\frac{2}{r_{12}} \deriv{j(r_{12},\mu)}{r_{12}}{} + \frac{1}{r_{12}} & = \frac{\text{erf}(\mu r_{12})}{r_{12}}, 
 \end{aligned}
\end{equation}
which is equivalent to 
\begin{equation}
 \label{def_j_0}
 \deriv{j(r_{12},\mu)}{r_{12}}{} = \frac{1 - \text{erf}(\mu r_{12})}{2}.
\end{equation}
The solution to Eq. \eqref{def_j_0} is 
\begin{equation}
 \label{eq:def_j}
 j(r_{12};\mu) = \frac{1}{2}r_{12}\bigg( 1 - \text{erf}(\mu r_{12})  \bigg) - \frac{1}{2\sqrt{\pi}\mu}e^{-(r_{12}\mu)^2},
\end{equation}
which depends on a unique parameter $\mu$, 
and the associated transcorrelated Hamiltonians obtained with $j(r_{12};\mu)$ is also controlled by $\mu$
\begin{equation}
 \label{eq:def_ht_mu}
 \tilde{H}[\mu] = e^{-j(r_{12};\mu)} H e^{j(r_{12};\mu)}. 
\end{equation}
Therefore, the right eigenvectors of $\tilde{H}[\mu]$ depends only on $\mu$, and the ground state eigenvector is defined by 
\begin{equation}
 \tilde{H}[\mu] \phimu(\br{}_{1},\br{}_{2}) = E_0 \phimu(\br{}_{1},\br{}_{2}), 
\end{equation}
where $E_0$ is the exact ground state energy. 
As $\tilde{H}[\mu]$ originates from a similarity transformation, it conserves the same spectrum as $H$ as long as it is developed in a complete basis set.

\subsubsection{A few properties of $j(r_{12};\mu)$}
The new Jastrow $j(r_{12};\mu)$ factor is tuned by a unique parameter, $\mu$, which controls the value of the leading term of the effective interaction at $r_{12}=0$ (see Eq.\eqref{def_j_00}). 
To understand the physical content together with the mathematical properties of $j(r_{12};\mu)$, 
one must keep in mind that the complete correlating factor $\mathcal{J}(r_{12};\mu)$ is 
\begin{equation}
 \mathcal{J}(r_{12};\mu) = \text{exp}\bigg(j(r_{12};\mu)\bigg),
\end{equation}
and that the exact wave function $\psiex(\br{}_1,\br{}_2)$ is formally expressed as 
\begin{equation}
 \psiex(\br{}_1,\br{}_2) = \mathcal{J}(r_{12};\mu) \phimu(\br{}_{1},\br{}_{2}).  
\end{equation}
From Eq.\eqref{eq:def_j} we can notice that 
\begin{equation}
 \lim_{r_{12} \rightarrow \infty}j(r_{12};\mu) = 0,
\end{equation}
which means that 
\begin{equation}
 \lim_{r_{12} \rightarrow \infty}\mathcal{J}(r_{12};\mu) = 1,
\end{equation}
and therefore $j(r_{12};\mu)$ impacts only the small $r_{12}$ behaviour,  
which is expected since the equation determining $j(r_{12},\mu)$ is obtained from the small $r_{12}$ limit of the Schroedinger equation (see Eq.\eqref{def_j_00}).

>>>>

The paper is organized as follows. 
In the first section, inspired by the short-range behaviour of RS-DFT effective wave functions,  
we derive the analytical form of a new Jastrow factor tuned by a unique parameter $\mu$, and discuss the physical content of such a simple Jastrow factor. 
We show then that the corresponding transcorrelated Hamiltonian has a straightforward analytical structure, which is briefly reviewed as a function of the $\mu$ parameter. 
Then, we perform a numerical study of the behaviour of the ground state eigenpair of $\tilde{H}[\mu]$ for the helium atom: 
in the first part we analyze the speed of convergence of the ground state energy with respect to the basis set and the parameter $\mu$, 
then we study the behaviour of the right eigenvectors in real space and compare it with the numerically exact ground state wave function. 
We show that the rapid convergence of the total energies coincides with a rapid convergence of the right eigenvectors in real space, and how to recover a very good approximation of the exact ground state wave function. 
Based on such encouraging results, we study the ground state energy of the helium isoelectronic series from Li$^+$ to Ne$^{8+}$ and compare different schemes to obtain 
a value of $\mu$ which depends on the system and still provides a fast convergence of the total energies. 
Eventually, we study the H$_2$ molecule and compare different regimes of correlation.

\section{A new form of Jastrow factor for transcorrelated Hamiltonian}
This section is dedicated to the derivation of a new form of Jastrow factor mimicking at short range the correlation effects of RS-DFT. 
For the sake of simplicity, we derive the transcorrelated Hamiltonian for the helium atom for a general Jastrow factor depending on the inter electronic distance.  
Then we focus on the behaviour at small $r_{12}$ of different flavours of Hamiltonians 
and the corresponding discontinuity in their eigenvectors. 
We browse three different types of Hamiltonians: the usual physical Hamiltonian, the transcorrelated Hamiltonian together with the effective Hamiltonian used in RS-DFT.
Based on such consideration, we propose a new Jastrow factor mimicking RS-DFT Hamiltonian at small $r_{12}$. 
Then, we derive the analytical form of the corresponding new transcorrelated Hamiltonians $\tilde{H}[\mu]$ in the case of a general $N$-electron system, and study its physical content as a  function of the parameter $\mu$.

\subsection{A simple physical picture: the case of the helium atom}
\subsubsection{The transcorrelated Hamiltonian for helium atom}
Let us write the Hamiltonian of the helium atom in the $ r_{12} = \big|{\bf r}_1 - {\bf r}_2 \big|$ coordinate 
\begin{equation}
 H  = h_c + \frac{1}{r_{12}},
\end{equation}
where 
\begin{equation}
 \begin{aligned}
 \label{def:h_c}
 h_c = &-\frac{1}{2} \sum_{i=1}^2 \bigg(\deriv{}{r_i}{2} + \frac{2}{r_i} \deriv{}{r_i}{} + \frac{2 Z}{r_i}\bigg) \\
     &-\bigg( \deriv{}{r_{12}}{2} + \frac{2}{r_{12}} \deriv{}{r_{12}}{} \bigg) \\
     &-\bigg( \frac{\bd{r_1}}{r_1} \cdot \frac{\bd{r_{12}} }{r_{12}}  \deriv{}{r_1}{} + 
                \frac{\bd{r_2}}{r_2} \cdot \frac{\bd{r_{21}} }{r_{21}}  \deriv{}{r_2}{} \bigg),
 \end{aligned}
\end{equation}
\label{sec:he_j}
Now let us consider the transcorrelated Hamiltonian $\tilde{H}[u]$ obtained by a similarity transformation of the usual Hamiltonian by a general Jastrow factor $u(r_{12})$ depending only on $r_{12}$: 
\begin{equation}
 \label{eq:ht_0}
 \begin{aligned}
 \tilde{H}[u]&= e^{-u(r_{12})} H e^{u(r_{12})}.
 \end{aligned}
\end{equation}
The only additional terms arising in $\tilde{H}[u]$ with respect to $H$ are those coming from the action of the differential operator in $r_{12}$,
\begin{equation}
 \mathcal{T}[u] =  -e^{-u(r_{12})}\bigg( \deriv{}{r_{12}}{2} + \frac{2}{r_{12}} \deriv{}{r_{12}}{} \bigg)e^{u(r_{12})}.  
\end{equation}
After some derivation, one can notice that by defining the following operators 
\begin{equation}
 \label{eq:def_tt}
 \tilde{t}[u] = -2 \deriv{u(r_{12})}{r_{12}}{} \deriv{}{r_{12}}{},
\end{equation}
\begin{equation}
 \label{eq:def_wt}
 \tilde{W}[u] = -\frac{2}{r_{12}} \deriv{u(r_{12})}{r_{12}}{}  , 
\end{equation}
\begin{equation}
 \label{eq:def_wt}
 \tilde{w}[u] = -\deriv{u(r_{12})}{r_{12}}{2} - \bigg( \deriv{u(r_{12})}{r_{12}}{} \bigg)^2, 
\end{equation}
one can write the Similarity transformed Hamiltonian  as
\begin{equation}
 \label{eq:ht_general}
 \tilde{H}[u] = h_c + \frac{1}{r_{12}}  + \tilde{W}[u] + \tilde{w}[u] + \tilde{t}[u] ,
\end{equation}
Therefore, one can see from Eq. \eqref{eq:ht_general} that the two electrons experience a modified scalar potential given by $\frac{1}{r_{12}}  + \tilde{W}[u] + \tilde{w}[u]$, and an additional differential operator $\tilde{t}[u]$, 
the latter making $\tilde{H}[u]$ non-hermitian.

\subsubsection{$r_{12} \rightarrow 0$ limits for different forms of Hamiltonians}
\label{sec:cusp}
Having established the analytical form of the transcorrelated Hamiltonian $\tilde{H}[u]$ for a general Jastrow factor $u(r_{12})$ in the case of the helium atom, one can then study the behaviour of $\tilde{H}[u]$ when $r_{12}\rightarrow 0$ and compare it to two other types of Hamiltonian: the usual physical Hamiltonian and that entering the RS-DFT framework. 
The analysis of the $r_{12}\rightarrow 0$ behaviour leads to the different condition that the eigenvectors of such operators must fulfill. For the sake of simplicity of the notations we focus on the ground state of these operators but the arguments are general for any bounded eigenvectors.

In the case of the usual Hamiltonian, the exact ground state wave function $\psiex$ must satisfy the Schroedinger equation in real space 
\begin{equation}
 H \psiex(\br{}_1,\br{}_2) = E_0 \psiex(\br{}_1,\br{}_2)\quad \forall (\br{}_1,\br{}_2).
\end{equation}
When looking at $r_{12}\approx 0$, all terms multiplying $\frac{1}{r_{12}}$ in $H \psiex(\br{}_1,\br{}_2)$ must remain finite, which, according to Eq.\eqref{def:h_c}, translates into
\begin{equation}
 \label{eq:cusp_0}
 \lim_{r_{12}\rightarrow 0}\bigg( \frac{2}{r_{12}} \deriv{}{r_{12}}{} -\frac{1}{r_{12}}\bigg) \psiex(\bd{r_1},\bd{r_2})  = c,\quad c<\infty 
\end{equation}
or equivalently by multiplying Eq. \eqref{eq:cusp_0} by $r_{12}$
\begin{equation}
 \lim_{r_{12}\rightarrow 0}\bigg( 2 \deriv{\psiex(\bd{r_1},\bd{r_2})}{r_{12}}{} -\psiex(\bd{r_1},\bd{r_2}) \bigg) = 0, 
\end{equation}
which translates into the famous cusp condition for antiparallel spins of Kato\cite{Kat-CPAM-57}, 
\begin{equation}
 \deriv{\psiex(\bd{r_1},\bd{r_2})}{r_{12}}{}\Bigr|_{r_{12}=0} = \frac{1}{2} \psiex(r_{12}=0). 
\end{equation} 
For an excellent pedagogical and general derivation of the cusp conditions, see Ref.\onlinecite{Tew-JCP-08}.

Regarding now the similarity transformed Hamiltonians, the exact ground state eigenvector must also satisfy an eigenvalue equation in real-space,
\begin{equation}
 \label{eq:hpi0_ex}
 \tilde{H}[u] \phiex(\br{}_1,\br{}_2) = E_0 \phiex(\br{}_1,\br{}_2) \quad \forall (\br{}_1,\br{}_2).
\end{equation}
Similarly to Eq. \eqref{eq:cusp_0}, when looking at $r_{12}\approx 0$, the terms multiplying $\frac{1}{r_{12}}$ in $\tilde{H}[u]\phiex(\br{}_1,\br{}_2)$ must remain finite, 
which involves also the term $\tilde{W}[u]$ in $\tilde{H}[u]$ (see Eq.\eqref{eq:def_wt}). Therefore, imposing the finiteness of the limit $r_{12}\rightarrow 0$ of Eq.\eqref{eq:hpi0_ex} translates into
\begin{equation}
 \label{eq:cusp_phi_0}
 \lim_{r_{12}\rightarrow 0}\bigg( \frac{2}{r_{12}} \deriv{}{r_{12}}{} -\frac{1}{r_{12}}\bigg[ 1 - 2 \deriv{u(r_{12})}{r_{12}}{}\bigg]\bigg) \phiex(\bd{r_1},\bd{r_2})  = c,\quad c<\infty.
\end{equation}
Then, the term $\frac{1}{r_{12}}\bigg[ 1 - 2\deriv{u(r_{12})}{r_{12}}{}\bigg]$ in Eq.\eqref{eq:cusp_phi_0} looks like an effective electron electron interaction induced by the presence of the Jastrow factor in $\tilde{H}[u]$. 
Therefore, as long as one imposes that 
\begin{equation}
 \label{eq:cusp_phi_1}
  \deriv{u(r_{12})}{r_{12}}{}\bigg|_{r_{12}=0} = \frac{1}{2},
\end{equation}
which is nothing but the cusp condition for the Jastrow factor $u(r_{12})$, the scalar term proportional to $\frac{1}{r_{12}}$ in Eq.\eqref{eq:cusp_phi_0} vanishes at $r_{12}=0$, and then one obtains a non-divergent effective electron-electron interaction. 
Within the condition of Eq.\eqref{eq:cusp_phi_1}, the general condition of Eq.\eqref{eq:cusp_phi_0} becomes 
\begin{equation}
 \label{eq:cusp_phi_2}
 \deriv{\phiex(\bd{r_1},\bd{r_2})}{r_{12}}{}\Bigr|_{r_{12}=0} = 0, 
\end{equation}
which implies that as long as the Jastrow factor contains the cusp conditions, the eigenfunction of $\tilde{H}[u]$ is cuspless. 
For instance, a simple Jastrow factor of the form $u(r_{12}) = \frac{1}{2} r_{12}$ releases $\phiex(\br{}_1,\br{}_2)$ from 
the constraint of fulfilling the cusp condition.

Another form of effective Hamiltonians leading to cuspless eigenfunctions are those obtained from RS-DFT 
where the Coulomb interaction $1/r_{12}$ is split into a non-divergent long-range interaction $\text{erf}(\mu r_{12})/r_{12}$ and a complementary short-range interaction $\text{erfc}(\mu r_{12})/r_{12}$, where $\text{erf}(x)$ and $\text{erfc}(x)$ are the error and complementary error functions, respectively. 
The parameter which controls such a splitting is the so-called range separation parameter $\mu$:  RS-DFT reduces the usual WFT when $\mu \rightarrow \infty$, and $\mu \rightarrow 0$ it gives back the usual Kohn-Sham theory.  
In practice, RS-DFT introduces a self-consistent Schroedinger-like equation which must be fulfilled by an effective wave function  $\Psi^\mu$.
Applied to a two electron system and looking in regions where $r_{12}\approx 0$, the leading terms of the self-consistent Schroedinger-like equation of RS-DFT reads 
\begin{equation}
 \label{eq:cusp_psi_mu_0}
 \lim_{r_{12}\rightarrow 0}\bigg( \frac{2}{r_{12}} \deriv{}{r_{12}}{} -\frac{\text{erf}(\mu r_{12})}{r_{12}}\bigg) \Psi^\mu(\bd{r_1},\bd{r_2})  = c,\quad c<\infty,  
\end{equation}
and as the effective interaction $\frac{\text{erf}(\mu r_{12})}{r_{12}}$ is non divergent at $r_{12}=0$ for finite values of $\mu$
\begin{equation}
 \lim_{r_{12} \rightarrow 0} \frac{\text{erf}(\mu r_{12})}{r_{12}} = \frac{2 \mu}{\sqrt{\pi}} , 
\end{equation}
one obtains that 
\begin{equation}
 \label{eq:cusp_psi_mu_1}
 \deriv{\Psi^\mu(\bd{r_1},\bd{r_2})}{r_{12}}{}\Bigr|_{r_{12}=0} = 0, 
\end{equation}
which means that the wave functions $\Psi^\mu$, just as $\phiex$ are cuspless, because they deal with an effective non-divergent interaction.

@@

\subsection{A new form of Jastrow factor $u(\mu,r_{12})$ mimicking a long-range interaction}
In section \ref{sec:cusp} we reviewed the forms of limit conditions at $r_{12}=0$ involved in different eigenvalue equations, which of course depends on the nature of the interaction at $r_{12}=0$. 
In that section we derive a new Jastrow factor inspired by the limit condition of the similarity transformed Hamiltonian (see Eq.\eqref{eq:cusp_phi_0}) which mimics that of RS-DFT (see Eq.\eqref{eq:cusp_psi_mu_0}), at least at leading order in $1/r_{12}$. 
\subsubsection{The working equation for $u(\mu,r_{12})$ }
Eq. \eqref{eq:ht_general} gives the similarity transformed Hamiltonian for an arbitrary Jastrow factor, 
and with respect to the usual Hamiltonian, $\tilde{H}[u]$ contains two new classes of terms: 
a new electron-electron potential $1/r_{12} + \tilde{w}[u]+ \tilde{W}[u]$ and a new differential operator  $\tilde{t}[u]$. 
Of course, these new terms depends on the explicit form of the Jastrow factor $u(r_{12})$ chosen. 
Now, we impose the form of $u(r_{12})$ such that it mimics, at leading order in $1/r_{12}$, the long-range effective interaction of RS-DFT. 
Mathematically, this conditions implies that 
\begin{equation}
 \begin{aligned}
 \label{def_j_00}
 \tilde{W}[u] + \frac{1}{r_{12}}&= \frac{\text{erf}(\mu r_{12})}{r_{12}} \\ 
\Leftrightarrow -\frac{2}{r_{12}} \deriv{u(r_{12},\mu)}{r_{12}}{} + \frac{1}{r_{12}} & = \frac{\text{erf}(\mu r_{12})}{r_{12}}, 
 \end{aligned}
\end{equation}
which is equivalent to 
\begin{equation}
 \label{def_j_0}
 \deriv{u(r_{12},\mu)}{r_{12}}{} = \frac{1 - \text{erf}(\mu r_{12})}{2}.
\end{equation}
The solution to Eq. \eqref{def_j_0} is 
\begin{equation}
 \label{eq:def_j}
 u(r_{12};\mu) = \frac{1}{2}r_{12}\bigg( 1 - \text{erf}(\mu r_{12})  \bigg) - \frac{1}{2\sqrt{\pi}\mu}e^{-(r_{12}\mu)^2},
\end{equation}
which depends on a unique parameter $\mu$, 
and the associated transcorrelated Hamiltonians obtained with $u(r_{12};\mu)$ is also controlled by $\mu$
\begin{equation}
 \label{eq:def_ht_mu}
 \tilde{H}[\mu] = e^{-u(r_{12};\mu)} H e^{u(r_{12};\mu)}. 
\end{equation}
Therefore, the right eigenvectors of $\tilde{H}[\mu]$ depends only on $\mu$, and the ground state eigenvector is defined by 
\begin{equation}
 \tilde{H}[\mu] \phimu(\br{}_{1},\br{}_{2}) = E_0 \phimu(\br{}_{1},\br{}_{2}), 
\end{equation}
where $E_0$ is the exact ground state energy. 
As $\tilde{H}[\mu]$ originates from a similarity transformation, it conserves the same spectrum as $H$ as long as it is developed in a complete basis set.

\subsubsection{A few properties of $u(r_{12};\mu)$}
The new Jastrow $u(r_{12};\mu)$ factor is tuned by a unique parameter, $\mu$, which controls the value of the leading term of the effective interaction at $r_{12}=0$ (see Eq.\eqref{def_j_00}). 
To understand the physical content together with the mathematical properties of $u(r_{12};\mu)$, 
one must keep in mind that the complete correlating factor $\mathcal{J}(r_{12};\mu)$ is 
\begin{equation}
 \mathcal{J}(r_{12};\mu) = \text{exp}\big(u(r_{12};\mu)\big),
\end{equation}
and that the exact wave function $\psiex(\br{}_1,\br{}_2)$ is formally expressed as 
\begin{equation}
 \psiex(\br{}_1,\br{}_2) = \mathcal{J}(r_{12};\mu) \phimu(\br{}_{1},\br{}_{2}).  
\end{equation}
From Eq.\eqref{eq:def_j} we can notice that 
\begin{equation}
 \lim_{r_{12} \rightarrow \infty}u(r_{12};\mu) = 0,
\end{equation}
which means that 
\begin{equation}
 \lim_{r_{12} \rightarrow \infty}\mathcal{J}(r_{12};\mu) = 1,
\end{equation}
and therefore $u(r_{12};\mu)$ impacts only the small $r_{12}$ behaviour,  
which is expected since the equation determining $u(r_{12},\mu)$ is obtained from the small $r_{12}$ limit of the Schroedinger equation (see Eq.\eqref{def_j_00}).
